Towards Detecting Stalkerware Through Network Traces

### 2019-03-23 paper/main.tex
Read-through up to Table 1

@@

As previously mentioned, spyware has a high turnover rate on the Google Play 
store, which leads to a practice we refer to as "reskinning:" a single app will 
be posted under many names, icons, and manufacturer names. Presumably, 
manufacturers do this because spyware gets reported frequently and taken down - 
we have already seen one instance of a spyware app getting removed from the 
Play store while we were examining it. Overt stalkerware which is not available 
through Google Play has also turned out to use reskinning, for reasons we 
haven't yet discovered. Apps can frequently be identified as copies of each 
other by observing their dashboards (where users view the communications and 
locations of the target phones), download sites, and advertisements. Several 
groups of off-store apps have nearly identical websites, in terms of 
organization, color, fonts, placement of buttons, prices, and other factors. 
For example, the "ILFMobile group" consists of three different companies 
(ILFMobile, PowerLine Group, and PhoneSpector) that claim to make a total of 
four different apps. Further investigation showed that all three companies had 
the same physical address in addition to identical websites. Unfortunately, 
they also had identical bugs in their payment methods, so we were unable to 
purchase the apps and verify that their network traces were identical. Other 
app sites appear to be similar to each other in other ways, 
with many having a large panel with a darkened image background extolling the 
features of the app above two buttons, one for a "live demo" and the other to 
"buy now." This is somewhat circumstantial evidence - it is entirely possible 
that these manufacturers simply chose similar web templates, or copied each 
other's site in an attempt to undercut competition. We therefore confirmed our 
suspicions by checking if the traffic signatures were also similar. These 
reskinned programs are potentially a boon to 
our investigation as the identical network signatures make it more likely that 
one signature generalizes across multiple apps. We have identified the 
following groups of applications that appear to be reskinned versions of each 
other.
\begin{enumerate}
	\item \textbf{SpyToMobile group}: Employee Work Spy by Piter Cline, SMS 
	Tracker by TheHar, Mobile Tracking by Antwat, Cell Phone Tracker by Susdel, 
	and Texting, Chat, 
	Phone Spy by MaxLo. All of these apps download the same program, which is 
	called SpyToMobile. Any app from this group installed after the first 
	states ``SpyToMobile is already installed.''
	\item \textbf{ILFMobile group}: HighsterMobile, PhoneSpector, Auto Forward 
	Spy, Surepoint Spy, and EasySpy. We were unable to purchase these and 
	ensure that they have the same signatures due to a bug in their payment 
	system that prevented us from purchasing them.
	\item \textbf{Codero group}: HelloSpy and TheTruthSpy have identical 
	network signatures as well as nearly identical websites.
	\item \textbf{Spyzie group}: Spyzie and SpyMyFone also have identical network signatures 
	as well as similar websites.
	\item \textbf{CallSMSTracker group}: Call and Message Tracker by Remote 
	Karath and Cell Tracker by Trackme. These apps have identical dashboards 
	and contact the same domains.
\end{enumerate}
We cannot guarantee that we have identified all of the members of any 
particular group. If we have not, it will make other members easier to 
identify, since signatures we have already found will scale to identify more 
apps than we had time to download and manually examine.

@@

\section{Related Work}
Previous studies [1,2], the media [3,4], and intimate partner violence (IPV) 
survivors have all reported that spyware is becoming an increasing source of 
intimate partner surveillance (IPS). Tracking functions supported by spyware 
allow stalkers to monitor communications, track locations, and remotely 
activate cameras and microphones [3,4]. In Dr. Ristenpart's work [5], they 
proposed methods for searching IPS-related apps. By Google search query 
expansion and unsupervised classification, they collected 9,224 apps possibly 
related to IPS on Google play and open-web. They manually selected 70 apps for 
investigation. They observed the most fundamental functions of these apps are 
monitoring locations, communication logs and data, media contents, and phone 
usages. They also found that existing anti-spyware tools cannot effectively 
detect IPS-related apps.
Mobile apps are often identified via the User-Agent string of the HTTP request, 
when one is available [7]. With the increased use of HTTPS this may not always 
be the case, but we expect to find (based on conversations with authors of 
previous work) that security and best coding practices are not prioritized by 
stalkerware developers. Malware identification may be of limited use as well, 
since techniques are available to cluster malware that comes from the same 
source code into "families." [8]

\section{Methodology}
\subsection{Finding Apps}
We selected thirty apps from a list provided by the team of Cornell researchers 
who performed the original large scale study of 
stalkerware\cite{chatterjee_spyware_2018}. An 
additional fifteen apps were included as they were recently added to the Google 
Play store. We found these apps by searching for keywords such as 
"spy","track", "locate", or "find". Encouragingly, IPV-related search terms 
such as "track girlfriend" return no results on the Play store. On the other 
hand, that did not mean stalking apps were not available on the play store, it 
just meant that different search terms had to be used to find them. Apps were 
chosen for their diversity of purpose and the probability that 
they can be used as stalkerware. These apps are the most likely to be 
downloaded when an abuser without deep technical knowledge tries to install 
stalkerware since they are either the first result upon searching for 
stalkerware, or are well known spyware apps due to advertisements.

We recognize that there are thousands of apps on the market that can be 
repurposed as stalkerware. Therefore, any useful process for identifying 
network signatures will eventually have to cover far more apps than we are 
starting out with. Our mere forty-five apps barely scratch the surface of the 
ecosystem, so they certainly do not provide a representative sample. Therefore, 
we did not employ any vigorous app selection methods, since we know this is 
just a starting point. Our hope is to eventually find a way to partially 
automate the signature identification process.

\subsection{Experimental Setup}
Out of the 45 apps we originally selected from Dr. Ristenpart's ML-curated list 
of stalkerware, 36 still existed on the app store. We installed these one at a 
time on one or both of two test phones, a Nexus 5X and a Nexus 6P.These models 
were chosen simply because they were readily available. Stalkerware user 
models  fall into two types: either an app is installed on the target phone and 
a login to a web portal is provided for the attacker to view the results, or 
the app comes with an ``attacker'' app and a ``target'' app, and the attacker 
app must be installed on another phone.

After installing an app, we first confirmed that it was functional and capable 
of acting like stalkerware. 
We then began tracking the phone's network traffic using Wireshark. For the 
majority of the apps, a network trace 
of five to ten minutes is enough to detect traffic going to and from the 
stalkerware app as long as the app's tracked functions like text or GPS are 
altered or the tracker app or website has a manual update function. Those that 
do not have such functionality have all been fake stalkerware apps. The network 
captures themselves contain at least one packet that provides easy manual 
detection of the related app.

@@

We created two aliases in order to avoid exposing any personal information to 
an ecosystem that is notorious for mishandling it\cite{ristenpart_ucsd_talk}. 
It is likely that the privacy of victims is 
not a high priority for anyone prepared to monetize abusive relationships, and 
stalkerware manufacturers have leaked victims' information in the past 
\cite{koebler_stalkerware_2018}. We took precautions by deleting all personal 
data that had been on the phones at the start of the project. However, we were 
not quite cautious enough - the app SpyMyFone managed to scrape a list of 
deleted contacts off of the Nexus 5X. As of this moment, we do not know how. We 
have deleted the account and requested that SpyMyFone delete all information 
associated with it from their records, but we have no evidence that they have 
complied. This served as a valuable and sobering lesson for the future.

The original goal of this project was to identify traffic signatures so that 
they could be detected by an IDS (Intrusion Detection System) or a network 
tap in an anonymized fashion. We hoped to use differential privacy techniques 
such as Google's RAPPOR algorithm \cite{erlingsson_rappor:_2014} to collect 
aggregated statistics about stalkerware use, without giving ourselves the 
ability to identify any particular device that had stalkerware installed. Upon 
further investigation before we began the project, this turned out to be 
insufficient protection. We as a computer 
scientists do not have the expertise required to assist victims of abusive 
relationships or domestic violence. We chose not to take our research in this 
direction to avoid causing any unintended harm. We limit ourselves to 
identifying traffic signatures on apps that we installed on personal phones.

Wiretapping the phones was accomplished by setting up our personal computers as 
WiFi hotspots. Since we controlled the passwords required to connect to our 
networks, and Windows alerts the user when a device connects to its mobile 
hotspot, we were able to ensure that we did not tap any traffic other than that 
of the test phones. 
\section{Results}
\subsection{App Traffic}

What causes most stalkerware to send packets? Our captures and testing show 
that there are three main categories of updating tracker state: manual, timed, 
and on tracked feature. Most free apps fall into the manual update method, 
where there is a button on a website or on a phone app that will send a signal 
packet to the stalkerware app that initializes the upload. Most paid apps also 
include a timed auto-update feature, where an update will be sent every few 
minutes. A very small subset of both advertised that it receives a "real time 
notification" every time email, texts, calls, or GPS, was changed or added to. 
Our results show that none of the stalkerware apps were able to send in real 
time, all the apps advertised as real time simply updated at a more frequent 
frequency than other apps.

Where do these stalkerware apps send their traffic? In general, all of our apps tested connected to a third party server that aggregated the information for display in either a website or a similar app for a separate phone. We had anticipated that at least one app would send tracking data directly to the abuser's phone, however this was not the case in testing. Perhaps setting up such a connection would be too difficult for someone who initiates intimate partner violence, or simply because it is easier to collect phone data in a dedicated server and use it for other purposes as well. Either way, all of the apps we found connected to a specific domain that was not the abuser's phone. This was both good and bad news. While it is much easier to find network signatures for packets being sent to a single destination with its own domain name, it also means that we will not be able to find an abuser through our method. This would have been an interesting tangent, unfortunately it does not work.
\begin{table*}
	\begin{tabular}{p{5cm}p{5cm}p{5cm}}
		\toprule
		Generalized Signature & Unique Apps & Total Apps \\
		\hline
		Domain contains "spy," "track," "find" & 6 & 12 \\
		SNI contains "spy," "track," "find" & 6 & 6 \\
		User-Agent contains "spy," "track," "find" & 1 & 3 \\
		User-Agent contains known app name & 1 & 1 \\
		Domain contains known app name & 3 & 4 \\
		SNI contains known app name & 1 & 1 \\
		SNI contains known app manufacturer & 3 & 3 \\
		Domain contains known app manufacturer & 3 & 3 \\
		\midrule
	\end{tabular}
\end{table*}
\subsection{Network Signatures}
\label{network_signatures}

When the network traffic is captured and the suspicious packets are identified, the next step would be to form a Network signature that can filter out those packets. But what can we use that isolates stalkerware traffic from normal traffic? To begin with, we must look at the packet identifiers that allow us to find the suspicious packets. 
\begin{itemize}
	\item Application Layer
	\begin{itemize}
		\item HTTP
		\begin{itemize}
			\item packet content
			\item User agent
		\end{itemize}
		\item HTTP and HTTPS headers
		\item Packet length
		\item Domain Name
		\item Host Name
		\item Location Requests
	\end{itemize}
	\item Transport Layer
	\begin{itemize}
		\item TLS 1.2
		\begin{itemize}
			\item Server Name Indication
		\end{itemize}
		\item TCP
		\begin{itemize}
			\item Ports
			\item Sequence number
		\end{itemize}
	\end{itemize}
	\item Internet Layer
	\begin{itemize}
		\item IP address
		\item IP version
	\end{itemize}
	\item Other
	\begin{itemize}
		\item Packet frequency
	\end{itemize}
\end{itemize}

All of the above signatures are able to be filtered out through a Wireshark filter or made into a rule for a Network Intrusion Detection System such as Snort. There are many other parts of the packets that we could attempt to match signatures for, and we will explain why we cannot add more to this list in a later section.

@@

The majority of stalkerware apps communicate though HTTP or HTTPS, while some 
use their own protocols, such as STUN. We examined the domain, SNI, User-Agent, and HTTP 
host, for each app and tried to isolate some part that we believed would be unique enough to identify that traffic. For the majority of the apps, they have a unique domain that each app connects to, which allowed for an easy method of filtering out their traffic. Some were hosted on shared cloud servers such as AWS, and so do not have a unique domain. These, however had a phrase in user agent or in the SNI that could be used for filtering.

Initially we attempted to add on our filters with an additional check for location requests, as many stalkerware apps will contain features for tracking location. Additionally, when connected to a network a phone may use the router to find its current location. Unfortunately, while we could occasionally see a external service being used, the majority of location requests were not listed in our network captures. That being said, our resultant domain, user agent, and SNI filters managed to achieve full coverage of all stalkerware apps.

 Unfortunately while this does catch all of the stalkerware apps we have found, it does not catch most of the stalkerware apps we have not found. In this case, we need to generalize our results into a format that can be either catch-all or easily scalable without requiring a network capture.

\subsection{Generalizing Signatures}

The words "spy," "find," and "track" appeared frequently in 
various fields of the traffic. The most common place to find them was in the 
destination domain, which we found using Wireshark's reverse DNS lookup tool. 
12 of the 26 apps that we eventually found signatures for contained these 
spyware keywords in their domains. We identified the following generalized 
signatures that can be used to identify most of our apps given nothing but 
information available on the Google Play store.

For the apps whose traffic do not contain "spy," "find," or "track," we will 
next check for the app name and manufacturer name or substrings of these two. 
Using these indicators has a significant disadvantage - we must be certain that 
the app is stalkerware and that the manufacturer doesn't make anything except 
stalkerware. In the case of smaller apps and reskinned apps, this may actually 
be a reasonable assumption - manufacturer names are generated to fool the 
Google Play store into thinking the app is new, not to identify a real company, 
and are therefore unique by necessity. We note that the two apps we identify 
only by manufacturer are made by GPSWOX and GeoZilla. GPSWOX makes 
location-tracking apps of all kinds, some of which may be difficult to use as 
stalkerware, but GeoZilla's only apps are all designed to track a phone. Since 
only two apps were identified this way, the false positive rate should be 
investigated further.
We identified the following generalized signatures that can be used to identify 
most of our apps given nothing but information available on the Google Play 
store. The second step of our analysis will be to search for these general 
signatures in our personal phones' traffic captures. This will give a better 
idea of whether it is possible to use general signatures without generating a 
prohibitive number of false positive stalkerware detections.

\section{Real world Traffic Testing}

In addition to the above network captures, we also took twenty-four hours of 
network traffic of our own phones as a baseline for false positive rates. These 
captures also included periods of time that were deemed high risk for false 
positives as they emulated behavior that is most likely to be caught by our 
signatures. Although this may not be enough to fully emulate an entire network, 
it is enough to prove the false positive rates on any given phone.

Using our specific signatures, our filters did not catch any packet that could be used as stalkerware. This is to be expected as we are very sure that there are no stalkerware apps installed on our personal phones and the filters used are quite specific to the apps that we looked at. Again, these filters are very specific with their requirements and does not scale when we add more apps to our study. However, we can confirm that stalkerware traffic is unique, and we have a filter that very accurately catches stalkerware.

Using the generic filters, we unfortunately wrongly caught 2 TCP streams and 3 domains out of our total of 2869 streams and 1756 domains in our self network capture. In addition, these filters will only catch 21 out of the 24 apps that comprise the specific filter. This is an issue as it does infer that our generic filters are too generic. Upon looking at the specified domains, we see that these are most likely adware services that track user browsing data. Although there is a possibility that these can be considered malicious enough to add to our results, these are not stalkerware and as such count towards our false positive rate. Also while the percent of false positives is quite low, if we were to deploy this filter in an actual router, it would undoubtedly collect far more domains than we should be catching.

>>>>

\section{Related Work}
Previous studies \cite{stories_from_survivors, burke_using_2011}, the media 
\cite{shahani_smartphones_nodate, chatterjee_spyware_2018}, and intimate 
partner violence (IPV) survivors \cite{ristenpart_ucsd_talk} have all reported 
that spyware is becoming an increasingly utilized means of performing
intimate partner surveillance (IPS). Tracking functions supported by spyware 
allow stalkers to monitor communications, track locations, and remotely 
activate cameras and microphones \cite{shahani_smartphones_nodate, 
chatterjee_spyware_2018}. Chatterjee et al. propose methods for searching for 
IPS-capable apps. By Google search query expansion and unsupervised 
classification, they collected 9,224 apps possibly related to IPS on Google 
play and the open Internet. They manually selected 70 apps for investigation. 
They observed the most fundamental functions of these apps are 
monitoring locations, communication logs and data, media contents, and phone 
usages. They also found that existing anti-spyware tools cannot effectively 
detect IPS-related apps \cite{chatterjee_spyware_2018}.
Mobile apps are often identified via the User-Agent string of the HTTP request, 
when one is available \cite{xu_identifying_2011}. With the increased use of 
HTTPS this may not always 
be the case, but we expect to find (based on conversations with authors of 
previous work) that security and best coding practices are not prioritized by 
stalkerware developers. Malware identification may be of use as well, 
since techniques are available to cluster malware that comes from the same 
source code into "families." \cite{hutchison_firma_2013}

@@

\begin{figure*}
	\centering
	\begin{subfigure}{0.8\columnwidth}
		\centering
		\includegraphics[width=0.9\linewidth]{../images/hellospy.png}
		\caption{HelloSpy's dashboard (cut off at bottom)}
		\label{fig:hellospy}
	\end{subfigure}%
	\begin{subfigure}{0.8\columnwidth}
		\centering
		\includegraphics[width=0.9\linewidth]{../images/thetruthspy.png}
		\caption{TheTruthSpy's dashboard}
		\label{fig:truthspy}
	\end{subfigure}
	\caption{HelloSpy and TheTruthSpy have remarkably similar websites. Their 
	traffic signatures also turned out to be identical, leading us to believe 
	they are made by the same manufacturer.}
	\label{fig:hellospy_thetruthspy}
\end{figure*}

As previously mentioned, spyware has a high turnover rate on the Google Play 
store, which leads to a practice we refer to as "reskinning:" a single app will 
be posted under many names, icons, and manufacturer names. Presumably, 
manufacturers do this because spyware gets reported frequently and taken down - 
we have already seen one instance of a spyware app getting removed from the 
Play store while we were examining it. Overt stalkerware which is not available 
through Google Play has also turned out to use reskinning, for reasons we 
haven't yet discovered. Apps can frequently be identified as copies of each 
other by observing their dashboards (where users view the communications and 
locations of the target phones), download sites, and advertisements. Several 
groups of off-store apps have nearly identical websites, in terms of 
organization, color, fonts, placement of buttons, prices, and other factors. 
For example, the "ILFMobile group" consists of three different companies 
(ILFMobile, PowerLine Group, and PhoneSpector) that claim to make a total of 
four different apps. Their websites, however, were remarkably similar (Figure 
\ref{fig:ilfmobiles}). Further investigation showed that all three companies 
had the same physical address in addition to identical websites. Unfortunately, 
they also had identical bugs in their payment methods, so we were unable to 
purchase the apps and verify that their network traces were identical. Another 
example of stalkerware websites that appear to be strangely similar to each 
other in other ways is shown in Figure \ref{hellospy_thetruthspy}. This is 
admittedly somewhat circumstantial evidence - it is entirely possible 
that these manufacturers simply chose similar web templates, or copied each 
other's site in an attempt to undercut competition. We therefore confirmed our 
suspicions by checking if the traffic signatures were also similar, and found 
that they were. These reskinned programs are potentially a boon to 
our investigation as the identical network signatures make it more likely that 
one signature generalizes across multiple apps. We have identified the 
following groups of applications that appear to be reskinned versions of each 
other.
\begin{enumerate}
	\item \textbf{SpyToMobile group}: Employee Work Spy by Piter Cline, SMS 
	Tracker by TheHar, Mobile Tracking by Antwat, Cell Phone Tracker by Susdel, 
	and Texting, Chat, 
	Phone Spy by MaxLo. All of these apps download the same program, which is 
	called SpyToMobile. Any app from this group installed after the first 
	states ``SpyToMobile is already installed.''
	\item \textbf{ILFMobile group}: HighsterMobile, PhoneSpector, Auto Forward 
	Spy, Surepoint Spy, and EasySpy. We were unable to purchase these and 
	ensure that they have the same signatures due to a bug in their payment 
	system that prevented us from purchasing them.
	\item \textbf{Codero group}: HelloSpy and TheTruthSpy have identical 
	network signatures as well as nearly identical websites.
	\item \textbf{Spyzie group}: Spyzie and SpyMyFone also have identical network signatures 
	as well as similar websites.
	\item \textbf{CallSMSTracker group}: Call and Message Tracker by Remote 
	Karath and Cell Tracker by Trackme. These apps have identical dashboards 
	and contact the same domains.
\end{enumerate}
We cannot guarantee that we have identified all of the members of any 
particular group. If we have not, it will make other members easier to 
identify, since signatures we have already found will scale to identify more 
apps than we had time to download and manually examine.

@@

\section{Methodology}

\subsection{Finding Apps}
We selected thirty apps from a list provided by the team of Cornell researchers 
who performed the original large scale study of 
stalkerware\cite{chatterjee_spyware_2018}. An 
additional fifteen apps were included because they were recently added to the 
Google Play store, or to replace apps that were found not to function. We found 
these apps simply by searching for them using keywords such as "spy" or 
"track." Encouragingly, specifically IPV-related search terms 
such as "track girlfriend" return no results on the Play store. On the other 
hand, that did not mean stalking apps were not available on the play store, it 
just meant that different search terms had to be used to find them. We selected 
a diverse set of purposes, from explicit spyware to innocuous dual-use apps. We 
predict that the apps we chose are likely to be 
downloaded when an abuser without deep technical knowledge tries to install 
stalkerware, since they are either among the first results upon searching for 
stalkerware, or are well advertised as spyware.

We recognize that there are possibly hundreds of apps on the market that can be 
repurposed as stalkerware. Therefore, any useful process for identifying 
network signatures will eventually have to cover far more apps than we are 
starting out with. Our mere forty-five apps barely scratch the surface of the 
ecosystem, so they may not provide a representative sample. Therefore, 
we did not employ any vigorous app selection methods, since we know this is 
just a starting point. Our hope is to eventually find a way to partially 
automate the signature identification process.

\subsection{Experimental Setup}
Out of the 45 apps we searched for for, 36 still existed on the app store. We 
installed these one at a 
time on one or both of two test phones, a Nexus 5X and a Nexus 6P. These phones 
were chosen simply because they were readily available. Stalkerware user 
models  fall into two types: either 
a login to a web portal is provided for the attacker to view the data, or 
the app comes with an ``attacker'' app and a ``target'' app, and the attacker 
app must be installed on another phone.

After installing an app, we first confirmed that it was functional and capable 
of acting like stalkerware. 
We then began tracking the phone's network traffic using Wireshark. For the 
majority of the apps, a network trace 
of five to ten minutes is enough to detect traffic going to and from the 
stalkerware app as long as the functions the app tracks like text or GPS are 
in use or the tracker app has manual update capabilities. We then manually 
examined the traffic, isolated that which belonged to the app, and determined 
the signature.

@@

We created two aliases in order to avoid exposing any personal information to 
an ecosystem that is notorious for mishandling it\cite{ristenpart_ucsd_talk}. 
It is likely that the privacy of victims is 
not a high priority for anyone prepared to monetize abusive relationships, and 
stalkerware manufacturers have leaked victims' information in the past 
\cite{koebler_stalkerware_2018, franceschi-bicchierai_this_2019}. We took 
precautions by deleting all personal 
data that had been on the phones at the start of the project. However, we were 
not quite cautious enough - the app SpyMyFone managed to scrape a list of 
deleted contacts off of the Nexus 5X. As of this moment, we do not know how. We 
have deleted the account and requested that SpyMyFone delete all information 
associated with it from their records, but we have no evidence that they have 
complied. This served as a valuable and sobering lesson for the future.

The original goal of this project was to identify traffic signatures so that 
they could be detected by an IDS (Intrusion Detection System) or a network 
tap in an anonymized fashion. We hoped to use differential privacy techniques 
such as Google's RAPPOR algorithm \cite{erlingsson_rappor:_2014} to collect 
aggregated statistics about stalkerware use, while denying ourselves the 
ability to identify any particular device that had stalkerware installed. Upon 
further investigation before we began the project, this turned out to be 
insufficient protection. We do not have the expertise required to assist 
victims of abusive 
relationships or domestic violence. If we were to uncover evidence of 
stalkerware, it is unclear what we should or could do with that information. We 
chose not to take our research in this 
direction to avoid causing any unintended harm. We limit ourselves to 
identifying traffic signatures on apps that we installed on personal phones. We 
leave the question of deploying our signatures to later work, once we have 
developed partnerships with experts who can advise us.

Wiretapping the phones was accomplished by setting up our personal computers as 
WiFi hotspots. Since we controlled the passwords required to connect to our 
networks, and Windows alerts the user when a device connects to its mobile 
hotspot, we were able to ensure that we did not tap any traffic other than that 
of the test phones. 
\section{Results}

\subsection{Network Signatures}
\label{network_signatures}

We identified the following traffic features as candidates for network 
signatures.
\begin{itemize}
	\item Application Layer
	\begin{itemize}
		\item HTTP
		\begin{itemize}
			\item Packet content
			\item User-Agent
			\item Presence of unique header fields 
			\item Packet length
			\item URI
			\item Host Name
		\end{itemize}
		\item Presence of excessive number of location requests
	\end{itemize}
	\item Transport Layer
	\begin{itemize}
		\item TLS 1.2
		\begin{itemize}
			\item Server Name Indicator
		\end{itemize}
		\item TCP
		\begin{itemize}
			\item Ports
		\end{itemize}
	\end{itemize}
	\item Internet Layer
	\begin{itemize}
		\item IP address
		\item Reverse-DNS-lookup domain name
	\end{itemize}
	\item Other
	\begin{itemize}
		\item Packet frequency
	\end{itemize}
\end{itemize}

Some of these features did not prove to be easy to use as signatures.
For example, different types of stalkerware exhibit wide variance in how often 
they send updates back to their servers. Some apps advertise 
``real time'' tracking, where new texts are uploaded to the service as soon as 
they are sent by or arrive at the phone, and location is updated as soon as the 
phone moves. In our experiments, very few of the apps who claimed this ability 
actually delivered on their promise. Texts were uploaded anywhere between ten 
minutes to a day after they were sent. Some of the stalkerware dashboards 
included a button that would update or refresh the data, but this usually did 
not trigger traffic to be sent from the phone.

Another feature we failed to identify was the presence of unusually high 
numbers of location requests. Since phones can (if GPS is not available) get 
information about their location by seeing which WiFi AP they are connected to, 
we attempted to identify packets that were requesting location from the AP, but 
were unsuccessful.

All of the other features proved to be useful in identifying some subset of the 
apps we examined.

@@

The majority of stalkerware apps communicate though HTTP or HTTPS. A minority 
use less common protocols, such as the app that could turn a phone into a live 
video camera remotely, which used STUN. We examined the features listed above 
for each app and tried to isolate some subset of them that we believed would be 
unique enough to identify that traffic. For the majority of the apps, the 
domain provided by a reverse DNS lookup was sufficient. This supported our 
assumption that most stalkerware does not attempt to obfuscate its traffic. 
Even servers hosted on Amazon's AWS had unique domain names, which usually 
consisted of some string identifying the app prepended to a string identifying 
which AWS datacenter the app is hosted by. We found signatures for 24 out of 
the 45 apps we examined. Of the remainder, nine were removed from the Google 
Play store before we began the project, and twelve either didn't work as 
stalkerware or didn't work at all.

\subsection{Generalizing Signatures}
\begin{table*}
	\begin{tabular}{p{6cm}p{4cm}p{4cm}}
		\toprule
		Generalized Signature & Unique Apps & Total Apps \\
		\hline
		Domain contains "spy," "track," "find" & 6 & 12 \\
		SNI contains "spy," "track," "find" & 6 & 6 \\
		User-Agent contains "spy," "track," "find" & 1 & 3 \\
		User-Agent contains known app name & 1 & 1 \\
		Domain contains known app name & 3 & 4 \\
		SNI contains known app name & 1 & 1 \\
		SNI contains known app manufacturer & 3 & 3 \\
		Domain contains known app manufacturer & 3 & 3 \\
		\midrule
	\end{tabular}
	\caption{Generalized signatures, created by identifying the similarities 
		between specific signatures}
	\label{tab:gen}
\end{table*}
Since the 24 apps that we manually identified signatures for undoubtedly 
represent a small proportion of all of the stalkerware available, we next 
attempted to look for patterns our signatures had in common that would allow us 
to generalize them. To our surprise, this was relatively straightforward.  
For example, the words "spy," "find," and "track" appeared frequently in 
various fields of the traffic. The most common place to find them was in the 
destination domain, which we found using Wireshark's reverse DNS lookup tool. 
12 of the 24 apps that we eventually found signatures for contained these 
spyware keywords in their domains. We therefore labeled this signature the 
``keyword method.''

For the apps whose traffic do not contain "spy," "find," or "track," we 
next check for the app name, the manufacturer name, or substrings of either. 
Using these indicators has a significant disadvantage - we must be certain that 
the app is stalkerware and that the manufacturer doesn't make anything except 
stalkerware. In the case of smaller apps and reskinned apps, this may actually 
be a reasonable assumption. Manufacturer names are generated to fool the 
Google Play store into thinking the app is new, not to identify a real company, 
and are therefore unique by necessity. We note that the two apps we can only 
identify by manufacturer are made by GPSWOX and GeoZilla. GPSWOX makes 
location-tracking apps of all kinds, some of which may be difficult to use as 
stalkerware, but GeoZilla's only apps are all designed to track a phone. Since 
only two apps were identified this way, the false positive rate should be 
investigated further.

The generalized signatures shown in Table \ref{tab:gen} can be used to identify 
all but three (12.5\%) of our apps given nothing but information available on 
the Google Play store. This means that if we were able to reliably identify 
apps on the 
Play Store than can be used as stalkerware, it would be possible to identify 
the traffic of a large fraction of them without having to manually find their 
signatures.

\section{Evaluation}

We evaluated both specific and general signatures by taking twenty-four hours 
of network traffic captures on our own personal phones. Our goal was to 
determine a lower bound for our false positive rates. These 
captures also included periods of time that were deemed high risk for false 
positives as they emulated behavior that is most likely to be caught by our 
signatures. Given the relatively small amount of data we have to work with, our 
results should be considered extremely conservative lower bounds on the true 
false positive rates of our signatures. TODO: explain how we got false pos rate

Using the specific signatures, our false positive rate was zero. This is to be 
expected as we are very sure that there 
are no stalkerware apps installed on our personal phones and the filters used 
are quite specific to the apps that we looked at.

Using the generic filters, we identified two TCP streams and 3 domains out of 
our total of 2869 streams and 1756 domains in our self network capture. In 
addition, these filters will only catch 21 out of the 24 apps that comprise the 
specific filter. This is an issue as it does infer that our generic filters are 
too generic. Upon looking at the specified domains, we see that these are most 
likely adware services that track user browsing data. Although there is a 
possibility that these can be considered malicious enough to add to our 
results, these are not stalkerware and as such count towards our false positive 
rate. Also while the percent of false positives is quite low, if we were to 
deploy this filter in an actual router, it would undoubtedly collect far more 
domains than we should be catching.
